Marine species do not need visas: the transboundary nature of the world’s exploited marine species

### 2020-08-07 Scripts/Manuscript_Draft.Rmd
All documents submitted to journal for revissions responses

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Exclusive Economic Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits of shared marine species. Here, we combined species distributions and the spatial delineation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the world's identified fished species are transboundary (n = 633). Between 2005 and 2014, fisheries targeting these species within the global-EEZs caught in average 48 million tonnes representing an average of USD 77 billion in fishing revenue per year. Some countries received over 90% of their catch and economic benefits from a few shared resources. Our analysis suggests that catches from transboundary species appear **to be declining** more than those from non-transboundary species. Our study has direct implications **for managing** fisheries of transboundary species, highlighting the need for strengthened international cooperation.

@@

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors as well as the species' evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. A species can be widely distributed (cosmopolitan) or endemic [@Nelson:2016kk]. However, fisheries management is predicated on the definition of “stocks” that, in most cases, are delineated by human-made spatial boundaries while such delineation often does not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The delineation of Exclusive Economic Zones (EEZs) under Part V of the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, extended political boundaries **from 12 to 200 nautical miles.** The boundaries of EEZs cut across the distribution of many species, creating shared stocks between nations [@UnitedNations:1986tl; @osthagen:2020hd]. **Shared stocks can be classified as three non-exclusive categories; "transboundary" - stocks that cross the EEZs of two or more coastal states; "straddling" - stocks that, cross neighboring EEZs and the adjacent high seas; and "highly migratory" - stocks that cross non-neighboring EEZs and the high seas (mainly tunas) [@Munro:2004th]. Here, our study focuses on the "transboundary" and "straddling" species that are exploited by fisheries operating within EEZs.**

Theories and empirical experience have shown that fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve their stated objectives [@Munro:2002uf; @Englander:2019dq].**For instance,** climate-driven changes in species distributions have led to conflicts between nations that were driven, at least partly, by changes in the proportion of captures [@Munro:2002uf], quota allocation, and fishery newcomers [@Spijkers:2017ij]. Moreover, climate change is likely to exacerbate these types of conflict and presents new challenges for political relations between neighboring countries [@Pinsky:2018cb] and for fisheries management [@Miller:2013iv]. Therefore, having an accurate understanding of the distribution and scale of transboundary and straddling fish stocks and fisheries is important to inform their sustainable management, particularly under climate change.

@@

We overlaid the known distribution of 938 commercially valuable marine species that are responsible for an average of 96.5% of global EEZ catch reconstructed by the *Sea Around Us* at the species level between 2005 and 2014, and 280 EEZs of 198 coastal **countries (See methods**). We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units), due to a lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level **(or even as groups containing multiple species). For example, shrimps (e.g. Litopenaeus stylirostris or Farfantepenaeus californiensis) in the Pacific coast of Mexico [@DiarioOficialdelaFederacionDOF:2018wt] and hammerhead shark (*Sphyrna zygaena*) in Peru [@MAP:2017uf] include multiple populations that are exploited by each country’s fisheries but they are managed at the species level in both cases.** Also, recent research has suggested strong connectivity between fish stocks, even when separated by thousands of kilometers [@Ramesh:2019va] through larval dispersal and **adult migration e.g. fishing migrating between reefs in neighboring EEZs [Levin:2018in]. Such ecological connectivity further highlights to the importance of accounting for transboundary stocks using biogeographical information and making the results relevant from a fisheries management perspective**. For this analysis we only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g. we did not include the high seas). We rely on multiple data sources including occurrence, distribution models and catch data and only consider a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods).

@@

**We identified 633 transboundary exploited species in the world (67.5% of the 938 species analyzed), almost double the previous estimates [@Teh:2015gd].** Between 2005 and 2014, national fleets targeting these transboundary species within EEZs **caught** an **annual** average of 48.5 million tonnes, representing 82.3% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level. These catches generated **a yearly average of USD 77,591 million in fishing revenue (78.5% of the global fishing revenue) during the same time period.** These values are more than the previously estimated 34.2 million tonnes in catches and USD 40,187 million (**in 2019 value**) in fishing revenue from shared stocks [@Teh:2015gd].**When we re-estimated the transboundary species’ catches using catch data that are consistent with those used in previous studies (FAO global reported data in year 2006), the 633 transboundary species we identified here accounted for 40.4 million tonnes of catches i.e., 18% higher than previously estimates. Our results suggests that the participation of transboundary species in global catch and fishing revenue might have been previously underestimated due to an incomplete knowledge on the transboundary nature of marine fished species.** The 305 species that are considered non-transboundary (termed as ‘discrete’ species here, see methods - Determining transboundary species trait) contributed a much smaller proportion of catches and revenues compared to those from the 633 transboundary species identified in this study. These discrete species represented only **an annual average of** 2.8 million tonnes and 4,282 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

In many cases, according to our categorization criteria, a transboundary species can distribute in multiple EEZs but only counted as transboundary in a subset of EEZs (see Methods). For example, the distribution of Peruvian anchoveta (*Engraulis ringens*) spans over the EEZs of Peru, Chile, and Ecuador. However, our study only counted the transboundary stocks in Peru and Chile [@Cashion:2018cg]. Anchoveta in Ecuadorian waters only include a small proportion of the shared distribution range **(spatial threshold between Ecuador and Peru < 10%; see methods - Criteria 3)** and thus do not meet our criteria to be considered as transboundary stock. Situation similar to the example of Peruvian anchoveta in Ecuadorian waters is common amongst the 633 transboundary species identified in this study. Overall, 590 of the 633 transboundary species have stocks in EEZs that are excluded from being counted as transboundary because they do not meet all our categorization criteria. The **annual average contribution from all the excluded stocks of transboundary species to fisheries (e.g., Ecuador’s anchoveta catch) between 2005 and 2014** was 10.8 million tonnes, representing USD 19,243 million in fishing revenue **over the same time period.**

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (**average country** revenue = 4,680 $\pm$ 6,000 million USD) and Eastern Asia (3,779 $\pm$ 3,093 million USD). The estimated **per country** revenues from transboundary species in these two regions are significantly different from other regions (one-way ANOVA; *DF*(16,165) = 5.081, *p* < 0.001, $\alpha$ = 0.05) (Fig. 1B). China (USD 7,284 million) and the USA (USD 11,604 million) contribute 55% and 82% to the **annual average** revenues in **Eastern** Asia and **Northern America**, respectively. In addition to China and the USA, Russia (USD 7,379 million), Peru (USD 6,044 **million**) and Japan (USD 3,907 **million**) are among the top five countries with the most fishing revenue generated from transboundary species between 2005 and 2014. These five nations were responsible for 41% of the yearly global fisheries revenue from transboundary species.

@@

![**Figure 1.** Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions: Northern America (N. America), Southern America (S. America), Latin America and the Caribbean (Ltn. Ame. & the Car.) Points = mean $\pm$ sd. Revenue in 2019 real USD.](../Figures/Fig1.png)

Peru and Russia, having the two largest fisheries by total catch in the world [@FAO:2018KK], caught over 5.8 million tonnes annually by each country on average between 2005 and 2014 (Supplementary Fig. S1). Peru's catches are mainly composed of Peruvian anchoveta (*Engraulis ringens*), shared with Chile, and account for 79% of the **national** transboundary species production. Recently these two countries have signed an agreement to work towards standardized stock assessments through coordinated management of this species [@UNDP:2016wp]. **Similarly, Russia, Japan and the USA have made a management agreement over Alaskan pollock (*Theragra chalcogramma*) in the Bering Sea since 1988 [@NOAA:gX6CtUTk].** Transboundary species also contribute largely to fisheries in **Eastern** Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). Particularly, China the world's top fish producer [@FAO:2018KK] **gets** one third (5.1 million tonnes) of the total catches from transboundary species, followed by Japan (1.8 million tonnes) and South Korea (1.06 million tonnes). Differences in transboundary fisheries can also be illustrated in terms of catch-revenue over the area of the EEZ $km^2$ (Fig. 2). As an example, the aggregated area of all EEZs of Northern European countries where transboundary species are present is 3.3 million $Km^2$, the $6^{th}$ smallest of the 17 groups analyzed (Supplementary Table S1). However, it has the second **highest average** revenue (USD 26.1 thousand per $Km^2$) and the **highest average*** catch (19.9 tonnes per $Km^2$) of transboundary species per EEZ area between 2005 and 2014. Countries from Western Europe get significantly more revenue from transboundary fisheries per $Km^2$ than from other regions (one-way ANOVA, *Df*(16,165) = 3.267, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.05).

@@

![Figure 2. Weighted benefits of transboundary species by $km^{2}$ and UN sub-region. A) Units in thousand 2019 USD. B) Units in tonnes. Points are sub-region averages $\pm$ s.d. by country](../Figures/Fig2.png)

**We determined the catch trend of each species within each EEZ (Fig. 3). While previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Therefore, our intention here is to examine where the catch trends of transboundary species are different from non-transboundary species between 2005 and 2010 relative to historic catch since 1951 (See methods).** Catch trends are classified as increasing (Category A), constant (Category B) or decreasing (Category C). We found significant differences within all catch trend categories for transboundary species (one-way ANOVA, *Df*(2,459) = 47.94, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.001) as opposite to discrete species where no significant differences where registered (one-way ANOVA, *Df*(2,106) = 1.885, *p* = 0.157; $\alpha$ = 0.05). We also found significant differences in catch trends when directly comparing transboundary to discrete species (MANOVA, *Df*(2,459) = 19.001, *p* < 0.001). **Overall, transboundary species that are only fished by one country are generally less likely to have a decreasing catch trend compared to when the species is shared by neighboring countries (Supplementary Table S2)**.

@@

![Figure 3. Number of transboundary species by catch trend and EEZ. Category A, Growing trend; Category B, Constant trend; Category C, Shrinking trend](../Figures/Fig3.png)

@@

Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the USA, Russia and the European Union (EU) [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches for Atlantic mackerel (*Scomber scombrus*) due to a climate-driven distribution shift [@Spijkers:2017ij]. Atlantic Mackerel is a transboundary species that **straddles into the high seas**. Amongst the countries involved in the 2007 Atlantic mackerel fisheries dispute, mackerel contributed an annual average (between 2005 and 2014) of 598.4 thousand tonnes (8.19%) to these countries’ total catches and USD 850.2 million (7.17%) to their total fishing revenue. Climate change is expected to continue changing the distribution and share proportion of fish stocks among countries up to a degree that new transboundary species will emerge [@Pinsky:2018cb], while some species will exit EEZs [@Oremus:2020id]. Exploring the detailed effects of climate change in the distribution of shared species is key to develop local adaptation methods to anticipate negative impacts to sustainability. **For example, understanding how climate change will modify the proportion of transboundary species that each neighboring EEZ share [@PalaciosAbrantes:2020vt], the time-frame at when the changes in sharing will happen, and the economic consequence of such effects can inform the development of more climate-resilient international treaties [@RashidSumaila:tp]**

@@

![Figure 4. Number of EEZs shared by transboundary species. A) Number of EEZs by species organized by associated habitat according to FishBase. Showing only species that share > 20 EEZs. B) Total catch between 2005 and 2014 of top five fishing countries capturing the top five shared species of each ecosystem preference (shown in bold on A). Note that there could be > 5 species due to similar sharing values. ** The category "Other" consists of species that have no classification according to FishBase.**](../Figures/Fig4.png)

**Many transboundary species are not considered highly migratory but still shared by numerous neighboring EEZs (e.g. Belone belone). In addition, many fish stocks have meta-populations that are connected through larvae dispersal with ‘source’ populations potentially supporting ‘sink’ populations thousands of kilometers away [@Ramesh:2019va]. For example, despite coral reef-related species being among the less shared neighboring EEZs identified in this study (Fig. 4), coral regions often share multiple species through larval connectivity [@Schill:2015fh] and adult movement [@Levin:2018in]. Species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm] but the potential boundary-crossing of these species must not be ignored. In some areas like the Caribbean and the Western Indian Ocean, transboundary marine protected areas have been identified as potential tools for achieving fisheries and conservation goals [@Levin:2018in; @Perez:2019dm]. The proper management of coral reef species is very important for many coastal communities, particularly in developing regions, that are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For instance some of the top countries fishing transboundary reef and pelagic-oceanic associated species (Fig. 4B**) are among the highest fish consumption countries [@FAO:2018KK]. In countries like the Philippines both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc].

@@

**The Sea Around Us has reconstructed global fisheries catches from 1950 to 2014 identifying commercial marine fish and invertebrates as well as fishing regions [@Zeller2016]. We used the Sea Around Us reconstructed species list to determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs. The total number of species analyzed was 938 representing 63.2% of identified species and accounting for 96.52% of the catch identified at the species' level.** To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (*i*) occurrence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (Table 1). Each source represents a different method of estimating the distribution of a given species and thus, provides a more robust result **than using only one data source**. Only commercial fished species with data from all four sources were included in the analysis.

@@

The previous models combine observational data with a series of biotic and abiotic data to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g., ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The *Sea Around Us* estimates total reconstructed catches - catches based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in available FAO data - for each country. Catches are also spatially allocated [@Zeller2016]. **Roughly, the Sea Around Us method takes each country's officially reported catch data (e.g. National, FAO or RFMO). Secondly, it uses literature (e.g. peer review, grey literature) to identify missing components (e.g. species, gears) and sources of alternative information for missing components. It then derives country estimates for missing data and creates time series interpolation. Finally the estimated and official data are aggregated for a total reconstructed catch data (See [@Pauly:2019bf] and [@Zeller2016] for catch reconstruction and spatial allocation details.)** We used the *Sea Around Us* catch reconstruction database from 2005 to 2014 as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs. We selected this time frame to investigate the recent (our time frame includes the **last decade of available data**) contribution of transboundary fisheries to catches and revenue from fisheries and to reduce the uncertainty embedded in the reconstruction process (See key uncertainties, below).

@@

**We used the occurrence database, the ENM-Nereus model, and SDM-SAU model to determine the presence of each species within each of the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells. All analyses only considered cases where all databases matched to obtain a more conservative estimate of transboundary species. Moreover, we assumed that a species was only present in a given grid-cell if it was reported in the Sea *Around Us catch* database. Therefore, all species that were not reported as caught in any single year between the reference years (2005 to 2014) in a given grid-cell were dropped. This rule assumes that if a commercial species is projected within any fishing country, such a species would have been fished, and thus reported at some point over the last decade of data (2005 - 2014), thus validating the models and only selected "current" shared species.**

@@

# Data Availability
The Sea Around Us datasets used are publicly available at http://www.seaaroundus.org/, the occurrence dataset is all publicly available (see Methods), the ENM-Nereus data can be accessed upon request to G.R. The final dataset and code are available at https://github.com/jepa/FishForVisa

>>>>

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Exclusive Economic Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits of shared marine species. Here, we combined species distributions and the spatial delineation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% **of the species analyzed** are transboundary (n = 633). Between 2005 and 2014, fisheries targeting these species within the global-EEZs caught in average 48 million tonnes representing an average of USD 77 billion in fishing revenue per year. Some countries received over 90% of their catch and economic benefits from a few shared resources. Our analysis suggests that catches from transboundary species appear **to be declining** more than those from non-transboundary species. Our study has direct implications **for managing** fisheries of transboundary species, highlighting the need for strengthened international cooperation.

@@

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors as well as the species' evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. A species can be widely distributed (cosmopolitan) or endemic [@Nelson:2016kk]. Fisheries management is predicated on the definition of “stocks” that, in most cases, are delineated by human-made spatial boundaries while such delineation often does not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The officialization of Exclusive Economic Zones (EEZs) under Part V of the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, extended political boundaries **from 12 to 200 nautical miles to give coastal nations property rights over marine resources [@osthagen:2020hd].** However, the boundaries of EEZs cut across the distribution of many species, creating shared stocks between nations [@UnitedNations:1986tl. **Shared stocks can be classified as three non-exclusive categories; "transboundary" - stocks that cross the EEZs of two or more coastal states; "straddling" - stocks that, cross neighboring EEZs and the adjacent high seas; and "highly migratory" - stocks that cross non-neighboring EEZs and the high seas (mainly tunas) [@Munro:2004th]. Here, our study focuses on the "transboundary" and "straddling" species that are exploited by fisheries operating within EEZs.**

Theories and empirical experience have shown that fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve their stated objectives [@Munro:2002uf; @Englander:2019dq]. **For instance,** climate-driven changes in species distributions have led to conflicts between nations that were driven, at least partly, by changes in the proportion of captures [@Munro:2002uf], quota allocation, and fishery newcomers [@Spijkers:2017ij]. Moreover, climate change is likely to exacerbate these types of conflict and presents new challenges for political relations between neighboring countries [@Pinsky:2018cb] and for fisheries management [@Miller:2013iv]. Therefore, having an accurate understanding of the distribution and scale of transboundary and straddling fish stocks and fisheries is important to inform their sustainable management, particularly under climate change.

@@

We overlaid the known distribution of 938 commercially valuable marine species that are responsible for an average of 96.5% of global EEZ catch reconstructed by the *Sea Around Us* at the species level between 2005 and 2014, and 280 EEZs of 198 coastal **countries (See methods**). We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units), due to a lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level **(or even as groups containing multiple species). For example, shrimps (e.g. Litopenaeus stylirostris or Farfantepenaeus californiensis) in the Pacific coast of Mexico [@DiarioOficialdelaFederacionDOF:2018wt] and hammerhead shark (*Sphyrna zygaena*) in Peru [@MAP:2017uf] include multiple populations that are exploited by each country’s fisheries, but they are managed at the species level in both cases.** Also, recent research has suggested strong connectivity between fish stocks, even when separated by thousands of kilometers [@Ramesh:2019va] through larval dispersal and **adult migration e.g. fishing migrating between reefs in neighboring EEZs [@Levin:2018in]. Such ecological connectivity further highlights to the importance of accounting for transboundary stocks using biogeographical information and making the results relevant from a fisheries management perspective**. For this analysis we only considered shared species between neighboring EEZs, rather than the species’ extended distribution (e.g. we did not include the high seas). We rely on multiple data sources including occurrence, distribution models and catch data and only consider a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods).

@@

**We identified 633 transboundary exploited species in the world (67.5% of the 938 species analyzed), almost double the previous estimates [@Teh:2015gd].** Between 2005 and 2014, national fleets targeting these transboundary species within EEZs **caught** an **annual** average of 48.5 million tonnes, representing 82.3% of the EEZ-derived catches reconstructed by the *Sea Around us* at the species level (Fig. 1A). These catches generated **a yearly average of USD 77,591 million in fishing revenue (78.5% of the global fishing revenue) during the same time period.** These values are more than the previously estimated 34.2 million tonnes in catches and USD 40,187 million (**in 2019 value**) in fishing revenue from shared stocks [@Teh:2015gd].**When we re-estimated the transboundary species’ catches using catch data that are consistent with those used in previous studies (FAO global reported data in year 2006), the 633 transboundary species we identified here accounted for 40.4 million tonnes of catches i.e., 18% higher than previously estimates. Our results suggests that the participation of transboundary species in global catch and fishing revenue might have been previously underestimated due to an incomplete knowledge on the transboundary nature of marine fished species.** The 305 species that are considered non-transboundary (termed as ‘discrete’ species here, see methods - Determining transboundary species trait) contributed a much smaller proportion of catches and revenues compared to those from the 633 transboundary species identified in this study. These discrete species represented only **an annual average of** 2.8 million tonnes and 4,282 million USD between 2005 and 2014. These results underscore the importance of transboundary species at the global level.

@@

![**Figure 1.** Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions: E. = Eastern, N. = Northern, S. = Southern, W. = Western, Ltn. Ame. & the Car. = Latin America and the Caribbean. Points = regional  mean ± sd. Revenue in 2019 real USD.](../Figures/Fig1.png)

In many cases, according to our categorization criteria, a transboundary species can distribute in multiple EEZs but only counted as transboundary in a subset of EEZs (see Methods). For example, the distribution of Peruvian anchoveta (*Engraulis ringens*) spans over the EEZs of Peru, Chile, and Ecuador. However, our study only counted the transboundary stocks in Peru and Chile [@Cashion:2018cg]. Anchoveta in Ecuadorian waters only include a small proportion of the shared distribution range **(spatial threshold between Ecuador and Peru < 10%; see methods - Criteria 3)** and thus do not meet our criteria to be considered as transboundary stock. Situation similar to the example of Peruvian anchoveta in Ecuadorian waters is common amongst the transboundary species identified in this study. Overall, 590 of the 633 transboundary species have stocks in EEZs that are excluded from being counted as transboundary because they do not meet all our categorization criteria. The **annual average contribution from all the excluded stocks of transboundary species to fisheries (e.g., Ecuador’s anchoveta catch) between 2005 and 2014** was 10.8 million tonnes, representing USD 19,243 million in fishing revenue **over the same time period.**

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America (**average country** revenue = 4,680 $\pm$ 6,000 million USD) and Eastern Asia (3,779 $\pm$ 3,093 million USD) (Fig. 1B). The estimated **per country** revenues from transboundary species in these two regions are significantly different from other regions (one-way ANOVA; *DF*(16,165) = 5.081, *p* < 0.001, $\alpha$ = 0.05). China (USD 7,284 million) and the USA (USD 11,604 million) contribute 55% and 82% to the **annual average** revenues from 2005 to 2014  in **Eastern** Asia and **Northern America**, respectively. In addition to China and the USA, Russia (USD 7,379 million), Peru (USD 6,044 **million**) and Japan (USD 3,907 **million**) are among the top five countries with the most fishing revenue generated from transboundary species between 2005 and 2014 (Fig. 1A). These five nations were responsible for 41% of the yearly global fisheries revenue from transboundary species.

Peru and Russia, having the two largest fisheries by total catch in the world [@FAO:2018KK], caught on average between 2005 and 2014 over 5.8 million tonnes by each country on average between 2005 and 2014 (Supplementary Fig. S1). Peru's catches are mainly composed of Peruvian anchoveta (*Engraulis ringens*), shared with Chile, and account for 79% of the **national** transboundary species production. Recently these two countries have signed an agreement to work towards standardized stock assessments through coordinated management of this species [@UNDP:2016wp]. **Similarly, Russia, Japan and the USA have made a management agreement over Alaskan pollock (*Theragra chalcogramma*) in the Bering Sea since 1988 [@NOAA:gX6CtUTk].** Transboundary species also contribute largely to fisheries in **Eastern** Asia (one-way ANOVA; *Df*(16,158) = 2.265, *p* = 0.005, $\alpha$ = 0.05). Particularly, China the world's top fish producer [@FAO:2018KK] **gets** one third (5.1 million tonnes) of the total catches from transboundary species, followed by Japan (1.8 million tonnes) and South Korea (1.06 million tonnes). Differences in transboundary fisheries can also be illustrated in terms of catch-revenue over the area ($km^2$) of the EEZ (Fig. 2). As an example, the aggregated area of all EEZs of Northern European countries where transboundary species are present is 3.3 million $km^2$, the $6^{th}$ smallest of the 17 groups analyzed (Supplementary Table S1). However, it has the second **highest average** revenue (USD 26.1 thousand per $km^2$) and the **highest average** catch (19.9 tonnes per $km^2$) of transboundary species per EEZ area between 2005 and 2014. Countries from Western Europe get significantly more revenue from transboundary fisheries per $km^2$ than any other country (one-way ANOVA, *Df*(16,165) = 3.267, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.05).

@@

![Figure 2. Weighted benefits of transboundary species by $km^{2}$ and UN sub-region. A) Revenue in thousand 2019 USD. B) Catch in tonnes. Points are sub-region averages $\pm$ s.d. by country](../Figures/Fig2.png)

**We determined the catch trend of each species within each EEZ classifying them as increasing (Category A), constant (Category B) or decreasing (Category C) (Fig. 3). While previous studies demonstrated that catch trends can be used to infer whether a stock is healthy, re-building, over-exploited or collapsed [@Kleisner:2011wn], several factors can influence stock status. Therefore, our intention here is to examine where the catch trends of transboundary species are different from non-transboundary species between 2005 and 2010 relative to historic catch since 1951 (See methods). We found significant differences within all catch trend categories for transboundary species (one-way ANOVA, *Df*(2,459) = 47.94, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* $\leq$ 0.001) as opposite to discrete species where no significant differences where registered (one-way ANOVA, *Df*(2,106) = 1.885, *p* = 0.157; $\alpha$ = 0.05). We also found significant differences in catch trends when directly comparing transboundary to discrete species categories (MANOVA, *Df*(2,459) = 19.001, *p* < 0.001). Overall, transboundary species that are only fished by one country are generally less likely to have a decreasing catch trend compared to when the species is shared by neighboring countries (Supplementary Table S2)**.

@@

![Figure 3. Number of transboundary species by catch trend and EEZ. Category A, Growing trend; Category B, Constant trend; Category C, Shrinking trend.“No Category” reflect species with less than 10 years of catch data and/ or less than 5 consecutive years of catch data.](../Figures/Fig3.png)

@@

Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the USA, Russia and the European Union (EU) [@Spijkers:2019cz]. For example, since 2007, the EU, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches for Atlantic mackerel (*Scomber scombrus*) due to a climate-driven distribution shift [@Spijkers:2017ij]. Atlantic Mackerel is a transboundary species that **straddles into the high seas**. Amongst the countries involved in the 2007 Atlantic mackerel fisheries dispute, mackerel contributed an annual average (between 2005 and 2014) of 598.4 thousand tonnes (8.19%) to these countries’ total catches and USD 850.2 million (7.17%) to their total fishing revenue. Climate change is expected to continue changing the distribution and share proportion of fish stocks among countries up to a degree that new transboundary species will emerge [@Pinsky:2018cb], while some species will exit EEZs [@Oremus:2020id]. Exploring the detailed effects of climate change in the distribution of shared species is key to develop local adaptation methods to anticipate negative impacts to sustainability. **. For example, understanding how climate change will modify the proportion of transboundary species that each neighboring EEZ share31, the time-frame at when the changes in sharing will happen, and the economic consequence of such effects can inform the development of more climate-resilient international treaties improving fisheries management [@RashidSumaila:tp]**

@@

![Figure 4. Number of EEZs shared by transboundary species. A) Number of EEZs by species organized by associated habitat according to FishBase. Showing only species that share > 20 EEZs. B) Average catch between 2005 and 2014 of top five fishing countries capturing the top five shared species of each ecosystem preference (shown in bold on a). Note that there could be > 5 species due to similar sharing values. ** The category "Other" consists of species that have no classification according to FishBase.**](../Figures/Fig4.png)

**Many transboundary species are not considered highly migratory but still shared by numerous neighboring EEZs (e.g. Belone belone) (Fig. 4A). In addition, many fish stocks have meta-populations that are connected through larvae dispersal with ‘source’ populations potentially supporting ‘sink’ populations thousands of kilometers away [@Ramesh:2019va]. For example, despite coral reef-related species being among the less shared neighboring EEZs identified in this study (Fig. 4A), coral regions often share multiple species through larval connectivity [@Schill:2015fh] and adult movement [@Levin:2018in]. Species with narrower ranges are often managed at the national level due to localized populations [@Payne:2008vm] but the potential boundary-crossing of these species must not be ignored. In some areas like the Caribbean and the Western Indian Ocean, transboundary marine protected areas have been identified as potential tools for achieving fisheries and conservation goals [@Levin:2018in; @Perez:2019dm]. The proper management of coral reef species is very important for many coastal communities, particularly in developing regions, that are highly dependent on subsistence fishing for food security and livelihoods [@CisnerosMontemayor:2016gq; @Wabnitz:2018gf]. For instance some of the top countries fishing transboundary reef and pelagic-oceanic associated species (Fig. 4B**) are among the highest fish consumption countries [@FAO:2018KK]. In countries like the Philippines both pelagic and reef fishes contribute substantially to both local food security and livelihood [@Cabral:2018hc].

@@

**The Sea Around Us has reconstructed global fisheries catches from 1950 to 2014 identifying commercial marine fish and invertebrates as well as fishing regions [@Zeller2016]. We used the Sea Around Us reconstructed species list to determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs. The total number of species analyzed was 938 representing 67% of identified species by the Sea Around Us and accounting for 96.52% of the catch identified at the species’ level.** To determine the current distribution of exploitable marine species, we used four data sources of species-distributions: (*i*) occurrence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (Table 1). Each source represents a different method of estimating the distribution of a given species and thus, provides a more robust result **than using only one data source**. Only commercial fished species with data from all four sources were included in the analysis.

@@

The previous models combine observational data with a series of biotic and abiotic data to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g., ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The *Sea Around Us* estimates total reconstructed catches - catches based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in available FAO data - for each country. Catches are also spatially allocated on a $0.5^{\circ}$ x $0.5^{\circ}$ latitude longitude grid [@Zeller2016]. **Roughly, the Sea Around Us method takes each country's officially reported catch data (e.g. National, FAO or RFMO). Secondly, it uses literature (e.g. peer review, grey literature) to identify missing components (e.g. species, gears) and sources of alternative information for missing components. It then derives country estimates for missing data and creates time series interpolation. Finally, the estimated and official data are aggregated for a total reconstructed catch data (See [@Pauly:2019bf] and [@Zeller2016] for catch reconstruction and spatial allocation details.)** We used the *Sea Around Us* catch reconstruction database from 2005 to 2014 as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs. We selected this time frame to investigate the recent (our time frame includes the **last decade of available data**) contribution of transboundary fisheries to catches and revenue from fisheries and to reduce the uncertainty embedded in the reconstruction process (See key uncertainties, below). Note that in call cases we report the average catch from 2005 to 2014.

@@

**We used the occurrence database, the ENM-Nereus model, and SDM-SAU model to determine the presence of each species within each of the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells. All analyses only considered cases where all databases matched to obtain a more conservative estimate of transboundary species. Moreover, we assumed that a species was only present in a given grid-cell if it was reported in the Sea *Around Us catch* database. Therefore, all species that were not reported as caught in any single year between the reference years (2005 to 2014) in a given grid-cell were dropped. This rule assumes that if a commercial species is projected within any fishing country, such a species would have been fished, and thus reported at some point over the last decade of data (2005 - 2014), thus validating the models and only selected "current" shared species. We acknowledge that this criteria might limit the global distribution of species thus resulting in a conservative estimate of transboundary species.**

@@

# Data Availability
The Sea Around Us datasets used are publicly available at http://www.seaaroundus.org/, the occurrence dataset is all publicly available (see Methods), the ENM-Nereus data can be accessed upon request to G.R. The processed results and code are available at https://github.com/jepa/FishForVisa
